---
title: "Predicting Heart Disease"
author: "Jesse Yan (netid@illinois.edu)"
date: "5/5/2021"
output:
  html_document: 
    theme: default
    toc: yes
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')
```

```{r, load-packages, include = FALSE}
# load packages
```

```{r read-full-data, warning = FALSE, message = FALSE}
# read full data
hd = readr::read_csv("data/hd.csv")

# clean data: remove variables missing over a third of data, then remove rows with missing data
na_prop = function(x) {
  mean(is.na(x))
}

hd_clean = na.omit(hd[, !sapply(hd, na_prop) > 0.33])
```

***

## Abstract

> Abstract text goes here.

***

## Introduction

Heart disease is one of the deadliest diseases in America. This analysis aims to help solve this issue by creating a tool that can screen for heart disease. The data for this project comes from the UCI Machine Learning Repository, collected from Cleveland Clinic Foundation; Hungarian Institute of Cardiology, Budapest; V.A. Medical Center, Long Beach, CA; and University Hospital, Zurich, Switzerland. 

***

## Methods

Describe the methods used in this section.

### Data

Specifically describe the data and how it is used here.

### Modeling

Discuss your modeling and validation strategy here.

***

## Results

State your results here.

***

## Discussion

Discuss your results here.

***

## Appendix

Place potential additional information here.
